---
title: "Analysis of 2017 General Election votes"
author: "Rajesh Korde"
output:
  html_document: default
  html_notebook: default
---

This documents analyzes the results of 2017 Primary Elections for Sammamish City Council. The data comes from [here](http://kingcounty.gov/depts/elections/results/2017/201711/reports.aspx.

```{r global_options, include=FALSE}
knitr::opts_chunk$set(cache=TRUE, fig.width=8, fig.height=6, echo=FALSE, 
                      warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(leaflet)
library(geojsonio)
library(forcats)
```

```{r}
precsp <- geojson_read("Data/Votes/json/wa.geojson", what = "sp")

raw2017 <- read_csv("Data/Votes/2017General.csv")

d <- raw2017

d <- mutate(d, CounterType = as.factor(CounterType)) %>%
  mutate(CounterType = fct_recode(CounterType,
           "Rituja" = "Rituja Indapure",
           "Chris" = "Chris Ross",
           "KarenM" = "Karen Moran",
           "KarenH" = "Karen N. Howe",
           "Pam" = "Pam Stuart",
           "John" = "John Robinson",
           "Jason" = "Jason Ritchie",
           "Mark" = "Mark Baughman",
           "Voters" = "Registered Voters")) %>%
  mutate(CounterType = fct_relevel(CounterType, "Rituja"))

voter_count <- filter(d, CounterType == "Voters") %>%
  group_by(Precinct) %>%
  mutate(RN = row_number()) %>%
  ungroup() %>%
  filter(RN == 1) %>%
  select(-RN)

d <- filter(d, CounterType != "Voters") %>%
  bind_rows(voter_count)
```

## Percent Vote count for Rituja

The map below shows the vote percentage (votes/Registered voters) for Rituja in each of the precints. Green indicates high vote percentage, red indicates low. Hovering on a precint will show the actual vote count and percentage.

```{r}
x1 <- filter(d, CounterType == "Rituja") %>%
  select(PRECNAME=Precinct, CounterType, SumOfCount)
precs <- precsp[precsp$PRECNAME %in% x1$PRECNAME,]
samm_precs <- x1$PRECNAME

samm_voters <- filter(d, CounterType == "Voters") %>%
  filter(Precinct %in% samm_precs) %>%
  select(PRECNAME=Precinct, CounterType, SumOfCount)

x1 <- bind_rows(samm_voters, x1) %>%
  spread(key=CounterType, value=SumOfCount) %>%
  mutate(VotePercent = (Rituja*100)/Voters)


pal <- colorNumeric("RdYlGn", domain = x1$VotePercent)
x2 <- sp::merge(x=precs, y=x1)

leaflet(x2) %>%
  addTiles() %>%
  addPolygons(stroke = TRUE, weight = 1,
              fillOpacity = 1,
              fillColor = ~pal(VotePercent),
              label = ~paste0(PRECNAME, " Vote Count: ", formatC(Rituja), " Vote Percent: ", formatC(VotePercent), "%"))

```


# Vote difference between Rituja and Chris

The plot below shows the vote difference normalized by number of registered voters in that precinct. So a vote differnce of 0.02 (shown when you hover over a precinct) means 2% more voters of that precinct voted for Rituja


```{r}
x1 <- filter(d, CounterType %in% c("Rituja", "Chris")) %>%
  filter(Precinct %in% samm_precs) %>%
  select(PRECNAME=Precinct, CounterType, SumOfCount) %>%
  bind_rows(samm_voters) %>%
  spread(key=CounterType, value=SumOfCount) %>%
  mutate(Diff = (Rituja - Chris)/Voters)

x2 <- sp::merge(x=precs, y=x1)
pal <- colorNumeric("RdYlGn", domain = x1$Diff)

leaflet(x2) %>%
  addTiles() %>%
  addPolygons(stroke = TRUE, weight = 1,
              fillOpacity = 1,
              fillColor = ~pal(Diff),
              label = ~paste0(PRECNAME, " Vote Difference: ", formatC(Diff)))
```

# Vote difference between Karen N Howe and Karen Moran

The plot below shows the vote difference normalized by number of registered voters in that precinct. So a vote differnce of 0.02 (shown when you hover over a precinct) means 2% more voters of that precinct voted for Karen H Howe.


```{r}
x1 <- filter(d, CounterType %in% c("KarenH", "KarenM")) %>%
  filter(Precinct %in% samm_precs) %>%
  select(PRECNAME=Precinct, CounterType, SumOfCount) %>%
  bind_rows(samm_voters) %>%
  spread(key=CounterType, value=SumOfCount) %>%
  mutate(Diff = (KarenH - KarenM)/Voters)

x2 <- sp::merge(x=precs, y=x1)
pal <- colorNumeric("RdYlGn", domain = x1$Diff)

leaflet(x2) %>%
  addTiles() %>%
  addPolygons(stroke = TRUE, weight = 1,
              fillOpacity = 1,
              fillColor = ~pal(Diff),
              label = ~paste0(PRECNAME, " Vote Difference: ", formatC(Diff)))
```


# Vote difference between Pam Stuart and John Robinson

The plot below shows the vote difference normalized by number of registered voters in that precinct. So a vote differnce of 0.02 (shown when you hover over a precinct) means 2% more voters of that precinct voted for Pam.


```{r}
x1 <- filter(d, CounterType %in% c("Pam", "John")) %>%
  filter(Precinct %in% samm_precs) %>%
  select(PRECNAME=Precinct, CounterType, SumOfCount) %>%
  bind_rows(samm_voters) %>%
  spread(key=CounterType, value=SumOfCount) %>%
  mutate(Diff = (Pam - John)/Voters)

x2 <- sp::merge(x=precs, y=x1)
pal <- colorNumeric("RdYlGn", domain = x1$Diff)

leaflet(x2) %>%
  addTiles() %>%
  addPolygons(stroke = TRUE, weight = 1,
              fillOpacity = 1,
              fillColor = ~pal(Diff),
              label = ~paste0(PRECNAME, " Vote Difference: ", formatC(Diff)))
```


# Vote difference between Mark Baughman and Jason Ritchie

The plot below shows the vote difference normalized by number of registered voters in that precinct. So a vote differnce of 0.02 (shown when you hover over a precinct) means 2% more voters of that precinct voted for Mark.

```{r}
x1 <- filter(d, CounterType %in% c("Mark", "Jason")) %>%
  filter(Precinct %in% samm_precs) %>%
  select(PRECNAME=Precinct, CounterType, SumOfCount) %>%
  bind_rows(samm_voters) %>%
  spread(key=CounterType, value=SumOfCount) %>%
  mutate(Diff = (Mark - Jason)/Voters)

x2 <- sp::merge(x=precs, y=x1)
pal <- colorNumeric("RdYlGn", domain = x1$Diff)

leaflet(x2) %>%
  addTiles() %>%
  addPolygons(stroke = TRUE, weight = 1,
              fillOpacity = 1,
              fillColor = ~pal(Diff),
              label = ~paste0(PRECNAME, " Vote Difference: ", formatC(Diff)))
```


# Table of votes for Rituja and Chris

Given below is the table of votes for Rituja and Chris by each precinct.


```{r}

x1 <- filter(d, CounterType %in% c("Rituja", "Chris")) %>%
  filter(Precinct %in% samm_precs) %>%
  select(PRECNAME=Precinct, CounterType, SumOfCount) %>%
  bind_rows(samm_voters) %>%
  spread(key=CounterType, value=SumOfCount)

library(kableExtra)
library(knitr)

kable(x1) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))


```

