---
title: "Kim/Manka 2018 Analysis"
author: "Rajesh Korde"
date: "December 26, 2018"
output: html_document
---

This documents analyzes the results of 2017 Primary Elections for Sammamish City Council. The data comes from [here](https://data.kingcounty.gov/Voting-Elections/Election-night-precinct-results-November-2018/m6sb-25v5).

```{r global_options, include=FALSE}
knitr::opts_chunk$set(cache=FALSE, fig.width=8, fig.height=6, echo=FALSE, 
                      warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(leaflet)
library(geojsonio)
library(forcats)
library(kableExtra)
library(knitr)
```

```{r}
precsp <- geojson_read("Data/Votes/json/wa.geojson", what = "sp")

raw2017 <- read_csv("Data/Votes/2017General.csv")
raw2018 <- read_csv("Data/Votes/2018General.csv")

raw2017 <- mutate(raw2017, Year = 2017)
raw2018 <- mutate(raw2018, Year = 2018)

d <- bind_rows(raw2018, raw2017)
d <- mutate(d, CounterType = as.factor(CounterType)) %>%
  mutate(CounterType = fct_recode(CounterType,
                                  "Rituja" = "Rituja Indapure",
                                  "Kim" = "Kim Schrier",
                                  "Manka" = "Manka Dhingra",
                                  "Voters" = "Registered Voters")) %>%
  mutate(CounterType = fct_relevel(CounterType, "Rituja"))

voter_count <- filter(d, CounterType == "Voters") %>%
  group_by(Precinct) %>%
  mutate(RN = row_number()) %>%
  ungroup() %>%
  filter(RN == 1) %>%
  select(-RN)

d <- filter(d, CounterType != "Voters") %>%
  bind_rows(voter_count)
```

## Percent Vote count for Rituja

The map below shows the vote percentage (votes/Registered voters) for Rituja in each of the precints. Green indicates high vote percentage, red indicates low. Hovering on a precint will show the actual vote count and percentage.

```{r}
x1 <- filter(d, CounterType == "Rituja") %>%
  select(PRECNAME=Precinct, CounterType, SumOfCount)
precs <- precsp[precsp$PRECNAME %in% x1$PRECNAME,]
samm_precs <- x1$PRECNAME

samm_voters <- filter(d, CounterType == "Voters") %>%
  filter(Precinct %in% samm_precs) %>%
  select(PRECNAME=Precinct, CounterType, SumOfCount)

x1 <- bind_rows(samm_voters, x1) %>%
  spread(key=CounterType, value=SumOfCount) %>%
  mutate(VotePercent = (Rituja*100)/Voters)


pal <- colorNumeric("RdYlGn", domain = x1$VotePercent)
x2 <- sp::merge(x=precs, y=x1)

leaflet(x2) %>%
  addTiles() %>%
  addPolygons(stroke = TRUE, weight = 1,
              fillOpacity = 1,
              fillColor = ~pal(VotePercent),
              label = ~paste0(PRECNAME, " Vote Count: ", formatC(Rituja), " Vote Percent: ", formatC(VotePercent), "%"))

```

# Comparison with Kim and Manka

In the maps below, green indicates the areas where Rituja did better than the opponent and Red indicates areas where the opponent performed much better. So in case of democratic candidates, Red areas are the best to go target.

### Kim Schrier

```{r}

plotDiff <- function(d, precs, name) {
  x1 <- filter(d, CounterType %in% c(name, "Rituja")) %>%
    select(PRECNAME=Precinct, CounterType, SumOfCount) %>%
    spread(key=CounterType, value=SumOfCount) %>%
    rename_(D1 = names(.)[2], D2 = names(.)[3]) %>%
    mutate(Diff = D1 - D2)
  
  x2 <- sp::merge(x=precs, y=x1)
  pal <- colorNumeric("RdYlGn", domain = x1$Diff)
  
  leaflet(x2) %>%
    addTiles() %>%
    addPolygons(stroke = TRUE, weight = 1,
                fillOpacity = 1,
                fillColor = ~pal(Diff),
                label = ~paste0(PRECNAME, " Vote Difference: ", formatC(Diff)))
}

plotDiff(d, precs, "Kim")
```

### Manka Dhingra 2017

```{r} 
plotDiff(filter(d, Year == 2017), precs, "Manka")
```

### Manka Dhingra 2018

```{r} 
plotDiff(filter(d, (CounterType == 'Rituja' & Year == 2017) | 
                  (CounterType == 'Manka' & Year == 2018)), precs, "Manka")
```

## Best precincts to target for campaigning

The table below shows the ranked list of precincts to go target based on the maximum vote difference between Rituja and kim/Manka2018. In the map, the red areas are the best areas to go campaign because they have the most vote difference.

```{r}
d1 <- filter(d, !((CounterType == 'Manka' & Year == 2017)))

x1 <- filter(d1, CounterType %in% c("Manka", "Kim", "Rituja")) %>%
  select(PRECNAME=Precinct, CounterType, SumOfCount) %>%
  filter(PRECNAME %in% samm_precs) %>%
  spread(key=CounterType, value=SumOfCount) %>%
  mutate(Max = ifelse(is.na(Manka), Kim, pmax(Manka, Kim))) %>%
  mutate(Diff = Max - Rituja) %>%
  arrange(desc(Diff))

kable(x1) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))


```
