---
title: "Analysis of 2017 Primary votes"
author: "Rajesh Korde"
output:
  html_document: default
  html_notebook: default
---

This documents analyzes the results of 2017 Primary Elections for Sammamish City Council. The data comes from [here](http://data.kingcounty.gov/Election-results).

```{r global_options, include=FALSE}
knitr::opts_chunk$set(cache=TRUE, fig.width=8, fig.height=6, echo=FALSE, 
                      warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(leaflet)
library(geojsonio)
library(forcats)
```

```{r}
precsp <- geojson_read("Data/Votes/json/wa.geojson", what = "sp")

raw2017 <- read_csv("Data/Votes/2017Primary.csv")
raw2016 <- read_csv("Data/Votes/2016General.csv")

d <- bind_rows(raw2017, raw2016)

d <- mutate(d, CounterType = as.factor(CounterType)) %>%
  mutate(CounterType = fct_recode(CounterType,
           "Rituja" = "Rituja Indapure",
           "Pam" = "Pam Stuart",
           "Karen" = "Karen N. Howe",
           "Ryika" = "Ryika Hooshangi",
           "Roger" = "Roger Goodman",
           "Minal" = "Minal Kode Ghassemieh",
           "Hillary" = "Hillary Clinton & Tim Kaine",
           "Trump" = "Donald J. Trump & Michael R. Pence",
           "Chris" = "Chris Ross",
           "Voters" = "Registered Voters")) %>%
  mutate(CounterType = fct_relevel(CounterType, "Rituja"))

voter_count <- filter(d, CounterType == "Voters") %>%
  group_by(Precinct) %>%
  mutate(RN = row_number()) %>%
  ungroup() %>%
  filter(RN == 1) %>%
  select(-RN)

d <- filter(d, CounterType != "Voters") %>%
  bind_rows(voter_count)
```

## Percent Vote count for Rituja

The map below shows the vote percentage (votes/Registered voters) for Rituja in each of the precints. Green indicates high vote percentage, red indicates low. Hovering on a precint will show the actual vote count and percentage.

```{r}
x1 <- filter(d, CounterType == "Rituja") %>%
  select(PRECNAME=Precinct, CounterType, SumOfCount)
precs <- precsp[precsp$PRECNAME %in% x1$PRECNAME,]
samm_precs <- x1$PRECNAME

samm_voters <- filter(d, CounterType == "Voters") %>%
  filter(Precinct %in% samm_precs) %>%
  select(PRECNAME=Precinct, CounterType, SumOfCount)

x1 <- bind_rows(samm_voters, x1) %>%
  spread(key=CounterType, value=SumOfCount) %>%
  mutate(VotePercent = (Rituja*100)/Voters)


pal <- colorNumeric("RdYlGn", domain = x1$VotePercent)
x2 <- sp::merge(x=precs, y=x1)

leaflet(x2) %>%
  addTiles() %>%
  addPolygons(stroke = TRUE, weight = 1,
              fillOpacity = 1,
              fillColor = ~pal(VotePercent),
              label = ~paste0(PRECNAME, " Vote Count: ", formatC(Rituja), " Vote Percent: ", formatC(VotePercent), "%"))

```


# Republican vs Democrat precincts

The map below shows relatively democrat vs republican precincts based on 2016 presidential elections. Note that Sammamish is pretty liberal and Hillary scored higher than Trump in every single precinct. Hovering on a precinct would show the raw vote count different. 

```{r}
x1 <- filter(d, CounterType %in% c("Trump", "Hillary")) %>%
  filter(Precinct %in% samm_precs) %>%
  select(PRECNAME=Precinct, CounterType, SumOfCount) %>%
  spread(key=CounterType, value=SumOfCount) %>%
  mutate(Diff = Hillary - Trump)
x2 <- sp::merge(x=precs, y=x1)
pal <- colorNumeric("RdYlBu", domain = x1$Diff)

leaflet(x2) %>%
  addTiles() %>%
  addPolygons(stroke = TRUE, weight = 1,
              fillOpacity = 1,
              fillColor = ~pal(Diff),
              label = ~paste0(PRECNAME, " Vote Difference: ", formatC(Diff)))
```

# Comparison with other 2017 candidates

In the maps below, green indicates the areas where Rituja did better than the opponent and Red indicates areas where the opponent performed much better. So in case of democratic candidates, Red areas are the best to go target.

### Karen N Howe

```{r}

plotDiff <- function(d, precs, name) {
  x1 <- filter(d, CounterType %in% c(name, "Rituja")) %>%
    select(PRECNAME=Precinct, CounterType, SumOfCount) %>%
    spread(key=CounterType, value=SumOfCount) %>%
    rename_(D1 = names(.)[2], D2 = names(.)[3]) %>%
    mutate(Diff = D1 - D2)
  
  x2 <- sp::merge(x=precs, y=x1)
  pal <- colorNumeric("RdYlGn", domain = x1$Diff)

  leaflet(x2) %>%
    addTiles() %>%
    addPolygons(stroke = TRUE, weight = 1,
                fillOpacity = 1,
                fillColor = ~pal(Diff),
                label = ~paste0(PRECNAME, " Vote Difference: ", formatC(Diff)))
}


plotDiff(d, precs, "Karen")
```

### Pam Stuart

```{r} 
plotDiff(d, precs, "Pam")
```

### Ryika Hoshangi

```{r} 
plotDiff(d, precs, "Ryika")
```


### Minal Ghassemieh

```{r}
plotDiff(d, precs, "Minal")
```

## Comparison with 2016 General Elections votes

In the maps below, green indicates the areas where Rituja did relatively better than the opponent and Red indicates areas where the opponent performed much better. So in case of democratic candidates, Red areas are the best to go target.

### Roger Goodman

```{r}
plotDiff(d, precs, "Roger")
```

### Hillary Clinton

```{r}
plotDiff(d, precs, "Hillary")
```



## Comparison of Votes from Rituja, Pam and Chris


The Chart below shows the relative vote diffences between Chris and Pam vs Chris and Rituja. The best precincts to target would be those with low vote difference between Chris and Pam and high vote difference between Chris and Rituja.

```{r fig.width=12, fig.height=8}
x1 <- filter(d, CounterType %in% c("Rituja", "Pam", "Chris")) %>%
  select(PRECNAME=Precinct, CounterType, SumOfCount)

x1 <- bind_rows(samm_voters, x1) %>%
  spread(key=CounterType, value=SumOfCount) %>%
  mutate(CPDiff = Chris - Pam) %>%
  mutate(CRDiff = Chris - Rituja)

norm <- filter(x1, PRECNAME != "SAM 45-3722") %>%
  mutate(CPDiff = if_else(CPDiff == 0, 0.5, as.numeric(CPDiff))) %>%
  mutate(temp = abs(CPDiff)) %>%
  mutate(norm = CRDiff/(11.2+temp)) %>%
  select(-temp)

library(ggrepel)
ggplot(x1, aes(x=CPDiff, y=CRDiff)) + 
  geom_point() +
  xlab("Chris-Pam Vote Diff") + ylab("Chris-Rituja Vote Diff") +
  geom_text_repel(aes(label=PRECNAME)) +
  theme_bw()
```


The plot below shows the best areas to target for campaigning. Green indicates small difference between Chris and Pam, but big difference between Chris and Rituja and red indicates big difference between Chris and Pam (or small difference between Chris and Rituja). Hovering should show differences between Chris/Pam and Chris/Rituja. 

```{r}
pal <- colorNumeric("RdYlGn", domain = norm$norm)
x2 <- sp::merge(x=precs, y=norm)


leaflet(x2) %>%
  addTiles() %>%
  addPolygons(stroke = TRUE, weight = 1,
              fillOpacity = 1,
              fillColor = ~pal(norm),
              label = ~paste0(PRECNAME, " Pam Vote Difference: ", formatC(CPDiff), " Rituja Vote Diff: ", formatC(CRDiff)))

```


Here is a list of precincts arranged from best to target to worst.

```{r}
library(kableExtra)
library(knitr)

tab <- rename(norm, "Total Voters" = Voters, 
              Precinct = PRECNAME,
              "Chris-Pam Diff" = CPDiff,
              "Chris-Rituja Diff" = CRDiff) %>%
  arrange(desc(norm)) %>%
  select(-c(norm))

kable(tab) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

## Best precincts to target for campaigning

The table below shows the ranked list of precincts to go target based on the maximum vote difference between Rituja and Pam/Karen. In the map, the red areas are the best areas to go campaign because they have the most vote difference.

```{r}


x1 <- filter(d, CounterType %in% c("Karen", "Pam", "Rituja")) %>%
  select(PRECNAME=Precinct, CounterType, SumOfCount) %>%
  spread(key=CounterType, value=SumOfCount) %>%
  mutate(Max = pmax(Karen, Pam)) %>%
  mutate(Diff = Rituja - Max)

x2 <- sp::merge(x=precs, y=x1)
pal <- colorNumeric("RdYlGn", domain = x1$Diff)

leaflet(x2) %>%
  addTiles() %>%
  addPolygons(stroke = TRUE, weight = 1,
              fillOpacity = 1,
              fillColor = ~pal(Diff),
              label = ~paste0(PRECNAME, " Vote Difference: ", formatC(Diff)))

tab <- mutate(x1, Diff = -Diff) %>%
  arrange(desc(Diff))

kable(tab) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))